The necessity of hypermedia RDF and an approach to achieve it

### 2012-04-12 eswc-2012/workshop2.tex
Add references for everything
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@219 43ef1190-90fd-4dd7-a769-d0121f02b80e

Mike Amundsen defines hypermedia types\footnote{\url{http://amundsen.com/hypermedia/}} as 
\begin{quote}
Hypermedia Types are MIME media types that contain native
hyper-linking semantics that induce application flow. For example,
HTML is a hypermedia type; XML is not.
\end{quote}
Furthermore, he defines a classification scheme called H~Factor as ``a
measurement of the level of hypermedia support and sophistication of a
media-type.'' The REST \& WOA Wiki defines ``the Hypermedia
Scale''\footnote{\url{http://restpatterns.org/Articles/The\_Hypermedia\_Scale}},
where the categorization is based on capabilities to do CRUD (Create,
Read, Update, Delete) operations.  Considered on their own, RDF
serializations are hypermedia types, but only at the \textsf{LO}
(outbound links) and \textsf{CL} (control data for links) levels (see
footnotes for details on this notation), and just an R~Type (read) on
the Hypermedia Scale. We aim at improving this situation.

Amundsen argues in a blog
post\footnote{\url{http://amundsen.com/blog/archives/1083}} that the
Semantic Web community should not pursue an API for RDF, but rather
make RDF serializations more powerful hypermedia types. He argues that
a key factor in the success of the Web is that messages not only
contain data but also application control information, and that this is
needed for the Web to scale.

@@

We see that with a simple vocabulary, RDF serializations can become a
very powerful hypermedia types. We also note that this vocabulary
enables agents to do the same operations as the SPARQL 1.1 Graph Store
HTTP
Protocol\footnote{\url{http://www.w3.org/TR/2011/WD-sparql11-http-rdf-update-20110512/}},
while being fully RESTful. The Protocol specification is currently not
RESTful per the discussion above as it requires extensive out-of-band
information, even though it was a key design goal. It should also be
possible to use this on the default (nameless) graph by assigning it a
name in the service description rather than defining it in an
out-of-band specification like is currently being done.

@@

  The RaUL\footnote{\url{http://vocab.deri.ie/raul}} goes beyond HTTP
  GET queries, which is what the \textsf{LT} factor is about, but can
  be used to formulate templated queries similar to those in HTML, or
  indeed to generate HTML fragments for the same purpose. This is also
  enables useful interactions, albeit quite different from those of
  this paper.

@@

In the Linked Data Design
Issue\footnote{\url{http://www.w3.org/DesignIssues/LinkedData}} Tim
Berners-Lee notes ``To make the data be effectively linked, someone
who only has the URI of something must be able to find their way the
SPARQL endpoint.'', but this is generally not possible today without
requiring out-of-band information.

The triples needed to do this are already defined in
VoID\footnote{\url{http://vocab.deri.ie/void}} and are in use:
\begin{verbatim}
<> void:inDataset [ void:sparqlEndpoint </sparql> . ] .
\end{verbatim}

@@

\bibliographystyle{plain-csmin}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
\bibliography{federation,webarch,webframeworks}

>>>>

Mike Amundsen defines hypermedia types\cite{hypermediatypes} as 
\begin{quote}
Hypermedia Types are MIME media types that contain native
hyper-linking semantics that induce application flow. For example,
HTML is a hypermedia type; XML is not.
\end{quote}
Furthermore, he defines a classification scheme called H~Factor as ``a
measurement of the level of hypermedia support and sophistication of a
media-type.'' The REST \& WOA Wiki defines ``the Hypermedia
Scale''\cite{hypermediascale},
where the categorization is based on capabilities to do CRUD (Create,
Read, Update, Delete) operations.  Considered on their own, RDF
serializations are hypermedia types, but only at the \textsf{LO}
(outbound links) and \textsf{CL} (control data for links) levels (see
\cite{hypermediatypes} for details on this notation), and just an R~Type (read) on
the Hypermedia Scale. We aim at improving this situation.

Amundsen argues in a blog
post\cite{nordfapi} that the
Semantic Web community should not pursue an API for RDF, but rather
make RDF serializations more powerful hypermedia types. He argues that
a key factor in the success of the Web is that messages not only
contain data but also application control information, and that this is
needed for the Web to scale.

@@

We see that with a simple vocabulary, RDF serializations can become a
very powerful hypermedia types. We also note that this vocabulary
enables agents to do the same operations as the SPARQL 1.1 Graph Store
HTTP Protocol\cite{graphstore}, while being fully RESTful. The
Protocol specification is currently not RESTful per the discussion
above as it requires extensive out-of-band information, even though it
was a key design goal. It should also be possible to use this on the
default (nameless) graph by assigning it a name in the service
description rather than defining it in an out-of-band specification
like is currently being done.

@@

  The RaUL\cite{raul} goes beyond HTTP GET queries, which is what the
  \textsf{LT} factor is about, but can be used to formulate templated
  queries similar to those in HTML, or indeed to generate HTML
  fragments for the same purpose. This is also enables useful
  interactions, albeit quite different from those of this paper.

@@

In the Linked Data Design
Issue\cite{linkeddataissue} Tim
Berners-Lee notes ``To make the data be effectively linked, someone
who only has the URI of something must be able to find their way the
SPARQL endpoint.'', but this is generally not possible today without
requiring out-of-band information.

The triples needed to do this are already defined in
VoID (see e.g. \cite{Alexander09describinglinked}) and are in use:
\begin{verbatim}
<> void:inDataset [ void:sparqlEndpoint </sparql> . ] .
\end{verbatim}

@@

\bibliographystyle{plain-csmin}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
\bibliography{federation,webarch,vocabs,specs,hypermedia}
